---
title: "Task 4 - Halloween Candy Data"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(janitor)
#install.packages("openxlsx")
library(openxlsx)
```

Initial look at the data

```{r}
raw_candy_2015 <- readxl::read_xlsx(here("raw_data/boing-boing-candy-2015.xlsx"))
raw_candy_2016 <- readxl::read_xlsx(here("raw_data/boing-boing-candy-2016.xlsx"))
raw_candy_2017 <- readxl::read_xlsx(here("raw_data/boing-boing-candy-2017.xlsx"))
```

steps: 
  remove unneccesary data (i.e anything not given a Joy, Despair, Meh score or specifically asked for in question)
     - for 2015, regex to get anything wrapped in "[]" 
     - for 2016, ^
     - for 2017, regex Q6 | 
"^[\\[][[:alnum:]]+[\\]]$"
```{r}
# This section shows my adventure in trying to work out why what i thought would be the same regex pattern produced different results

# only_candy_2015 <- raw_candy_2015 %>% 
#   select(matches("^[[[:print:]]+]$"))
# 
# whats_the_difference <- raw_candy_2015 %>% 
#   select(matches("^[[][[:print:]]+[]]$"))
# 
# whats_the_difference2 <- raw_candy_2015 %>% 
#   select(matches("^[[]{2}[[:print:]]+[]]{1}$"))
# 
# list_a <- names(only_candy_2015) 
# list_b <- names(whats_the_difference)
# 
# list_a[!(list_a %in% list_b)]
```
    
  
## 2015 DATA

# Separate candy data to reattach to relevant columns later    
```{r}
only_candy_2015 <- raw_candy_2015 %>% 
   select(matches("^[[]{1}[[:print:]]+[]]{1}$"))
```

```{r}
clean_names_candy_2015 <- janitor::clean_names(only_candy_2015)


```

# Separate age and trick or treating columns and a few lines to  mutate trick or treat into logical, just in case
also cleaning up the names 

```{r}
only_personal_data_2015 <- raw_candy_2015 %>% 
  select(`How old are you?`, `Are you going actually going trick or treating yourself?`) %>% 
  rename("age" = "How old are you?", "going_trick_or_treating" = "Are you going actually going trick or treating yourself?") %>% 
  mutate(year = 2015, .before = "age") %>% 
  mutate(gender = NA, .after = "age") %>% 
  mutate(country = NA, .after = "gender")
  

# only_personal_data_2015 %>% 
#   mutate(trick_or_treating_logical = case_when(
#     going_trick_or_treating == "No" ~ FALSE,
#     going_trick_or_treating == "Yes" ~ TRUE
#     
#   ))
```

rejoining data back together

```{r}
rejoined_2015 <- bind_cols(only_personal_data_2015, clean_names_candy_2015)

```

## 2016 DATA

# Separate candy data to reattach to relevant columns later  

```{r}
only_candy_2016 <- raw_candy_2016 %>% 
   select(matches("^[[]{1}[[:print:]]+[]]{1}$"))

clean_names_candy_2016 <- janitor::clean_names(only_candy_2016)

# Comparing column names 

# list_a <- names(clean_names_candy_2015)
# list_b <- names(clean_names_candy_2016)
# 
# tibble(list_a[!(list_a %in% list_b)])
# tibble(list_b[!(list_b %in% list_a)])

```


```{r}
# This converts changes the country column to a less stupid name and standardises the data 
# This will not pick up every single variant - However, it picks up most of them
country_tidied_2016 <- raw_candy_2016 %>% 
  select(`Which country do you live in?`) %>% 
  mutate(`Which country do you live in?` = tolower(`Which country do you live in?`)) %>% 
  rename(country = `Which country do you live in?`) %>% 
  mutate(country = case_when(
    country == NA  ~ "Other",
    str_detect(country, "^us|america|states$") ~ "USA",
    str_extract(country, "[u]{1}[s]{1}|[eu]+rica|trump") == TRUE ~ "USA",
    str_detect(country, "kingdom$|^uk|england|scotland|britain") ~ "UK",
    str_detect(country, "canada") ~ "Canada",
    TRUE ~ "Other"
  ))
```

```{r}
# extract relevant personal data so it can be tidied and reconnected
only_personal_data_2016 <- raw_candy_2016 %>% 
  select(`How old are you?`, `Your gender:`,`Are you going actually going trick or treating yourself?` ) %>% 
  rename("age" = "How old are you?", 
         "going_trick_or_treating" = "Are you going actually going trick or treating yourself?",
         "gender" = "Your gender:") %>% 
  mutate(year = 2016, .before = "age")
         
```

```{r}
rejoined_2016 <- bind_cols(only_personal_data_2016, country_tidied_2016) %>% 
  bind_cols(clean_names_candy_2016)
```


```{r}
# ## Testing reading country using regex
# raw_candy_2016 %>% 
#   select(`Which country do you live in?`) %>% 
#   mutate(`Which country do you live in?` = tolower(`Which country do you live in?`)) %>% 
#   rename(country = `Which country do you live in?`) %>% 
#   mutate(country = case_when(
#     country == NA  ~ "Other",
#     str_detect(country, "^us|america|states$|trump") ~ "USA",
#     str_extract(country, "[u]{1}[s]{1}|america") == TRUE ~ "USA",
#     str_detect(country, "kingdom$|^uk|england|scotland") ~ "UK",
#     str_detect(country, "canada") ~ "Canada",
#     TRUE ~ "Other"
#   ))
# 
# country_tidied_2016
```




## 2017 DATA

# Separate candy data to reattach to relevant columns later  

```{r}
only_candy_2017 <- raw_candy_2017 %>%
  select(matches("^Q6"))
#This pulls out all the "candy" data (assuming of course that all the relevant data is under Q6) - But they have "Q6 | " in the column names 
```

```{r}
# This removes the "Q6 | " from all the columns  
colnames(only_candy_2017) <- gsub("Q6 [\\|] ", " ",colnames(only_candy_2017))
```

```{r}
# Clean candy names
clean_names_candy_2017 <- janitor::clean_names(only_candy_2017) 

# comparing column names 
# 
# list_a <- names(clean_names_candy_2015)
# list_b <- names(clean_names_candy_2017)
# 
# tibble(list_a[!(list_a %in% list_b)])
# tibble(list_b[!(list_b %in% list_a)])
```


# remove personal data 
Like with 2016, country has been left at this stage to be cleaned separately
```{r}
only_personal_data_2017 <- raw_candy_2017 %>% 
  select(`Q3: AGE`, `Q2: GENDER`, `Q1: GOING OUT?`) %>% 
  rename("age" = "Q3: AGE", 
         "going_trick_or_treating" = "Q1: GOING OUT?",
         "gender" = "Q2: GENDER") %>% 
  mutate(year = 2017, .before = "age")
```

# cleaning the country data - as before, converting to lower case, then using regex patterns to standardise naming
```{r}
country_tidied_2017 <- raw_candy_2017 %>% 
  select(`Q4: COUNTRY`) %>% 
  mutate(`Q4: COUNTRY` = tolower(`Q4: COUNTRY`)) %>% 
  rename(country = `Q4: COUNTRY`) %>% 
  mutate(country = case_when(
    country == NA  ~ "Other",
    str_detect(country, "^us|america|states$") ~ "USA",
    str_extract(country, "[u]{1}[s]{1}|[eu]+rica|trump") == TRUE ~ "USA",
    str_detect(country, "kingdom$|^uk|england|scotland|britain") ~ "UK",
    str_detect(country, "canada") ~ "Canada",
    TRUE ~ "Other"
  ))
```

# rejoining 2017 data

```{r}
rejoined_2017 <- bind_cols(only_personal_data_2017, country_tidied_2017) %>% 
  bind_cols(clean_names_candy_2017)
```

## rejoining the datasets and final cleaning

```{r}
rejoined_candy <- full_join(rejoined_2015, rejoined_2016) %>%
  full_join(rejoined_2017)
```

```{r}
# check to see how many NA's were present in age before mutating to integer (317)
# rejoined_candy %>% 
#   select(age) %>% 
#   filter(is.na(age)) %>% 
#   count()
```


```{r}
# changing the age column to integer - this will eliminate any "non-number" answers (i.e. "enough" "fifty-four") and replace them with NA 
# I have decided to leave the NA's in the data
# Logically, age cannot be over a certain value - I've decided to replace anything over 100 with NA
rejoined_candy <- rejoined_candy %>% 
  mutate(age = as.integer(age)) %>% 
  mutate(age = case_when(
    age > 100 ~ NA,
    age <= 100 ~ age
  ))


# rejoined_candy %>%
#   filter(is.na(age)) %>%
#   count()
```

```{r}
#checking gender column
unique(rejoined_candy$gender)

# counting NA's to check next step has worked
# rejoined_candy %>%
#   select(gender) %>%
#   filter(is.na(gender)) %>%
#   count()


# replacing NA's with "Not Provided" 
rejoined_candy <- rejoined_candy %>% 
  mutate(gender = replace_na(gender, "Not Provided"))

# doing the same for country - This is technically not necessary, but i think makes it clearer why 2015 has no country data
rejoined_candy <- rejoined_candy %>% 
  mutate(country = replace_na(country, "Not Provided in 2015 data"))
```


# Writing cleaned data

```{r}
clean_candy <- rejoined_candy

write.xlsx(clean_candy, here("clean_data/clean_candy.xlsx"))
```

