---
title: "Task 4 - Halloween Candy Data - Analysis questions"
output: html_notebook
---


```{r}
library(tidyverse)
library(here)
library(openxlsx)
candy <- read.xlsx(here("clean_data/clean_candy.xlsx"))
```

What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
# this selects all candy columns, mutates them into a logical TRUE or FALSE depending on if they are NA or if they have values
# from there, sums all the values that are TRUE
# after, it sums those values into a new column - giving us the total amount of ratings
# The total ratings is 386924 out of 1131229 possible (9349*121) 

candy %>% 
  select(!c(year, age, gender, country, going_trick_or_treating)) %>% 
  mutate(across(.col = is.character, .fns = is.na)) %>% 
  summarise(across(.col = where(is.logical), .fns = sum)) %>%
  mutate(total_ratings = sum(across(where(is.integer)))) %>% 
  select(total_ratings)
```

What was the average age of people who are going out trick or treating?

```{r}
candy %>% 
  select(age, going_trick_or_treating) %>% 
  drop_na() %>% 
  filter(going_trick_or_treating == "Yes") %>% 
  mutate(avg_age_T_T = mean(age))
```


What was the average age of people who are not going trick or treating?
For each of joy, despair and meh, which candy bar received the most of these ratings?
How many people rated Starburst as despair?

For the next three questions, count despair as -1, joy as +1, and meh as 0.

What was the most popular candy bar by this rating system for each gender in the dataset ?
What was the most popular candy bar in each year?
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?