---
title: "Task 4 - Halloween Candy Data - Analysis questions"
output: html_notebook
---


```{r}
library(tidyverse)
library(here)
library(openxlsx)
candy <- read.xlsx(here("clean_data/clean_candy.xlsx"))
```

What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
# this selects all candy columns, mutates them into a logical TRUE or FALSE depending on if they are NA or if they have values
# from there, sums all the values that are TRUE
# after, it sums those values into a new column - giving us the total amount of ratings
# The total ratings is 386924 out of 1159276 possible (9349*124) - Around 33%

candy %>% 
  select(!c(year, age, gender, country, going_trick_or_treating)) %>% 
  mutate(across(.col = is.character, .fns = is.na)) %>% 
  summarise(across(.col = where(is.logical), .fns = sum)) %>%
  mutate(total_ratings = sum(across(where(is.integer)))) %>% 
  select(total_ratings)
```

What was the average age of people who are going out trick or treating?

```{r}
candy %>% 
  select(age, going_trick_or_treating) %>% 
  drop_na() %>% 
  filter(going_trick_or_treating == "Yes") %>% 
  mutate(avg_age_t_t = mean(age)) %>% 
  select(avg_age_t_t) %>% 
  head(1)
```


What was the average age of people who are not going trick or treating?

```{r}
candy %>% 
  select(age, going_trick_or_treating) %>% 
  drop_na() %>% 
  filter(going_trick_or_treating == "No") %>% 
  mutate(avg_age_no_t_t = mean(age)) %>% 
  select(avg_age_no_t_t) %>% 
  head(1)
```


For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
# for each candy, count "joy", "despair", "meh"
candy_types <- colnames(only_candy)
length_candy_types <- length(candy_types)
current_value <- 1


while(current_value < length_candy_types){
  select() %>%
  group_by(candy_types) %>%
  count() %>%
  slice(1)
  str_c()
  
  
  current_value <- only_candy[current_value + 1]
    }

```

```{r}
only_candy %>%
  select(butterfinger) %>%
  group_by(butterfinger) %>%
  count() %>%
  slice(1)

only_candy <- replace(only_candy,is.na(only_candy), "No Answer")
```

```{r}

## TOMORROW - TRY AND GET A LOOP WORKING TO COUNT JOY
for(candy_type in only_candy){
  select() %>% 
  group_by() %>% 
  count() %>% 
  slice(1)
}
```

```{r}
 only_candy <- candy %>% 
  select(-c(year, age, gender, country, going_trick_or_treating)) %>% 
  
  

```


How many people rated Starburst as despair?

For the next three questions, count despair as -1, joy as +1, and meh as 0.

What was the most popular candy bar by this rating system for each gender in the dataset ?
What was the most popular candy bar in each year?
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?