---
title: "Task 4 - Halloween Candy Data - Analysis questions"
output:
  html_document:
    df_print: paged
---


```{r}
library(tidyverse)
library(here)
library(openxlsx)
candy <- read.xlsx(here("clean_data/clean_candy.xlsx"))
```

## Cleaning

For cleaning this data, i went in with the mindset that anything that had a Joy, Despair or Meh rating was something that should be kept, even if it clearly 
wasnt candy. This also meant that anything that listed additional items, or anything that didn't have any real relevance to the tasks was dropped.
My logic for dropping the lists was that anything included in them was likely a one off or just nonsense that would add unnecessary columns or have very little 
impact on existing numbers.

My first step was to split the 2015 data into "Candy" and "Personal Info" -
Personal Info contained things required by the questions, such as age, gender, country,etc.
Candy was anything wrapped in square brackets, after a quick observation of the data.
To get this data, I used regex, using the following pattern:

```{r}
# "^[[]{1}[[:print:]]+[]]{1}$"
```

This searched for 1 opening square bracket at the start, multiple of anything that can be printed, then ending with 1 closing square bracket.
The column names output were then cleaned up using Janitor

I then created a separate dataset for just personal information - Which in 2015 consisted of only Age and Going Trick or Treating. 
To make life a bit easier, i mutated a few new columns in, as seen below

```{r}
# only_personal_data_2015 <- raw_candy_2015 %>% 
#   select(`How old are you?`, `Are you going actually going trick or treating yourself?`) %>% 
#   rename("age" = "How old are you?", "going_trick_or_treating" = "Are you going actually going trick or treating yourself?") %>% 
#   mutate(year = 2015, .before = "age") %>% 
#   mutate(gender = NA, .after = "age") %>% 
#   mutate(country = NA, .after = "gender")
```

I then joined the two datasets back together to create rejoined_2015 using bind_cols, as there were no common columns to use a Join

-----------------------------
2016 followed using the same approach.
However, i chose to separate the Country data from the rest of the personal info, as this required extra regex. 
This column was largely full of nonsense, so quite a lot of options were required.

```{r}
# # This converts changes the country column to a less stupid name and standardises names

# # This will not pick up every single variant - However, it picks up most of them
 
# country_tidied_2016 <- raw_candy_2016 %>% 
#   select(`Which country do you live in?`) %>% 
#   mutate(`Which country do you live in?` = tolower(`Which country do you live in?`)) %>% 
#   rename(country = `Which country do you live in?`) %>% 
#   mutate(country = case_when(
#     country == NA  ~ "Other",
#     str_detect(country, "^us|america|states$") ~ "USA",
#     str_extract(country, "[u]{1}[s]{1}|[eu]+rica|trump") == TRUE ~ "USA",
#     str_detect(country, "kingdom$|^uk|england|scotland|britain") ~ "UK",
#     str_detect(country, "canada") ~ "Canada",
#     TRUE ~ "Other"
#   ))
```

I decided not to worry too much about countries outside of what was specifically asked for, replacing them and all NA's with Other.

This was then reattached to the 2016 candy and personal info data sets.

--------------------------
2017 was, again, similar to both of the above, with some minor changes.

The regex used to pull out the candy data now looked like this:

```{r}
# #This pulls out all the "candy" data (assuming of course that all the relevant data is under Q6) - But they have "Q6 | " in the column names

# only_candy_2017 <- raw_candy_2017 %>%
#   select(matches("^Q6"))



# # This removes the "Q6 | " from all the columns  
# colnames(only_candy_2017) <- gsub("Q6 [\\|] ", " ",colnames(only_candy_2017))

```
 And had the additional step of using gsub and colnames to pull out the "Q6 | " from all the candy columns.
 
--------------------------
After following the same processes as above, I now had rejoined_2015, rejoined_2016 and rejoined_2017.

To combine all these, i used bind_cols on 2015 and 2016, then the same again on the newly bound 2015_2016 and 2017

--------------------------
## Final Steps of Cleaning

I then removed any NA's in the age, gender and country columns

```{r}
# # changing the age column to integer - this will eliminate any "non-number" answers (i.e. "enough" "fifty-four") and replace them with NA 
# # I have decided to leave the NA's in the data
# # Logically, age cannot be over a certain value - I've decided to replace anything over 100 with NA
# rejoined_candy <- rejoined_candy %>% 
#   mutate(age = as.integer(age)) %>% 
#   mutate(age = case_when(
#     age > 100 ~ NA,
#     age <= 100 ~ age
#   ))
```

```{r}
# # replacing NA's with "Not Provided" 
# rejoined_candy <- rejoined_candy %>% 
#   mutate(gender = replace_na(gender, "Not Provided"))
# 
# # doing the same for country - This is technically not necessary, but i think makes it clearer why 2015 has no country data
# rejoined_candy <- rejoined_candy %>% 
#   mutate(country = replace_na(country, "Not Provided in 2015 data"))
```


----------------------

# At this point, i felt the data was cleaned enough to be useable for analysis. 


---------------------

## Analysis


What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
# this selects all candy columns, mutates them into a logical TRUE or FALSE depending on if they are NA or if they have values
# from there, sums all the values that are TRUE
# after, it sums those values into a new column - giving us the total amount of ratings
# The total ratings is 386924 out of 1159276 possible (9349*124) - Around 33%

candy %>% 
  select(!c(year, age, gender, country, going_trick_or_treating)) %>% 
  mutate(across(.col = is.character, .fns = is.na)) %>% 
  summarise(across(.col = where(is.logical), .fns = sum)) %>%
  mutate(total_ratings = sum(across(where(is.integer)))) %>% 
  select(total_ratings)
```

What was the average age of people who are going out trick or treating?

```{r}
candy %>% 
  select(age, going_trick_or_treating) %>% 
  drop_na() %>% 
  filter(going_trick_or_treating == "Yes") %>% 
  mutate(avg_age_t_t = mean(age)) %>% 
  select(avg_age_t_t) %>% 
  head(1)
```


What was the average age of people who are not going trick or treating?

```{r}
candy %>% 
  select(age, going_trick_or_treating) %>% 
  drop_na() %>% 
  filter(going_trick_or_treating == "No") %>% 
  mutate(avg_age_no_t_t = mean(age)) %>% 
  select(avg_age_no_t_t) %>% 
  head(1)
```


For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
#making a subset of data containing only candy columns and replacing any NA's with characters "No Answer" 
# Incase there is any need to sum

only_candy <- candy %>% 
  select(-c(year, age, gender, country, going_trick_or_treating)) 

only_candy <- replace(only_candy,is.na(only_candy), "No Answer")

# This turns the candy columns and ratings into rows

tidyr::pivot_longer(only_candy, cols = everything(), names_to = "names", values_to = "rating") %>% 
  group_by(names, rating) %>% 
  arrange(names) %>% 
  count(rating) %>% 
  filter(rating != "No Answer") %>% 
  arrange(desc(n)) %>% 
  ungroup(names) %>% 
  slice(1)

# Some might argue that a broken glow stick is not candy - To satisfy any nitpickers, here is the answer if they are excluded
tidyr::pivot_longer(only_candy, cols = everything(), names_to = "names", values_to = "rating") %>% 
  group_by(names, rating) %>% 
  arrange(names) %>% 
  count(rating) %>% 
  filter(rating != "No Answer", names != "broken_glow_stick") %>% 
  arrange(desc(n)) %>% 
  ungroup(names) %>% 
  slice(1)


```

How many people rated Starburst as despair?

```{r}
# Using only_candy
only_candy %>% 
  select(starburst) %>% 
  mutate(starburst_despair = sum(starburst == "DESPAIR")) %>% 
  select(starburst_despair) %>% 
  head(1)
```


For the next three questions, count despair as -1, joy as +1, and meh as 0.
```{r}
# Separating code into two blocks - personal and candy data
# converting candy responses into numeric values then reattaching to personal data
# This creates a new data set to avoid future faffing about

only_personal <- candy %>% 
  select(year, age, gender, country, going_trick_or_treating)

candy_numeric <- only_candy %>%
  mutate(across(everything(), ~  case_when(
    . == "JOY" ~ 1,
    . == "DESPAIR" ~ -1,
    . == "MEH" ~ 0,
    . == "No Answer" ~ 0,
    TRUE ~ 0))) 

full_candy_numeric <- bind_cols(only_personal, candy_numeric)

```

What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
# This converts all the candy data into a new rows called "candy_name" and "rating" to allow the total rating to be counted


full_candy_numeric %>%
  select(!c(year, age, country, going_trick_or_treating)) %>% 
  tidyr::pivot_longer(where(is.double), names_to = "candy_name", values_to = "rating") %>% 
  group_by(gender, candy_name) %>% 
  summarise(across(is.double, sum)) %>% 
  slice_max(rating)

  
```

What was the most popular candy bar in each year?


```{r}
# Same as above, but with the added step of pivoting only columns that are not year 
full_candy_numeric %>%
  select(!c(age, gender, country, going_trick_or_treating)) %>% 
  tidyr::pivot_longer(cols = -year, names_to = "candy_name", values_to = "rating") %>% 
  group_by(year, candy_name) %>% 
  summarise(across(is.double, sum)) %>% 
  slice_max(rating)

```

What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# Same as above - This step could technically be improved, as during cleaning, i elected to replace anything that wasn't US, Canada, or UK with other
# If we wanted a more complete analysis for each country, we would leave other countries as they were.

full_candy_numeric %>%
  select(!c(year, age, gender, going_trick_or_treating)) %>% 
  tidyr::pivot_longer(where(is.double), names_to = "candy_name", values_to = "rating") %>% 
  group_by(country, candy_name) %>% 
  summarise(across(is.double, sum)) %>% 
  slice_max(rating)
```

